---
  title: "Social Network Analysis with igraph"
---

This tutorial was motivated by:

NetSciX School of Code Workshop
Network analysis with R and igraph
Wroclaw, Poland, January 10 2016

Katya Ognyanova, katya@ognyanova.net
www.kateto.net/netscix2016


```{r,eval=FALSE}
library(igraph)
library(igraphdata)
````


# Networks in igraph

```{r}
g1 <- graph( edges=c(1,2, 2,3, 3,1), n=3, directed=F ) # an undirected graph with 3 edges
# The numbers are interpreted as vertex IDs, so the edges are 1-->2, 2-->3, 3-->1
plot(g1) # A simple plot of the network - we'll talk more about plots later
class(g1)
g1
```

```{r}
g2 <- graph( edges=c(1,2, 2,3, 3,1), n=10 ) # now with 10 vertices, and directed by default
plot(g2)   
g2
```



```{r}
g3 <- graph( c("John", "Jim", "Jim", "Jill", "Jill", "John")) # named vertices
# When the edge list has vertex names, the number of nodes is not needed
plot(g3)
g3
```

```{r}
g4 <- graph( c("John", "Jim", "Jim", "Jack", "Jim", "Jack", "John", "John"), 
             isolates=c("Jesse", "Janis", "Jennifer", "Justin") )  
```


In named graphs we can specify isolates by providing a list of their names.

```{r}
plot(g4, edge.arrow.size=.5, vertex.color="gold", vertex.size=15, 
     vertex.frame.color="gray", vertex.label.color="black", 
     vertex.label.cex=1.5, vertex.label.dist=2, edge.curved=0.2) 
```


Small graphs can also be generated with a description of this kind:
'-' for undirected tie, "+-' or "-+" for directed ties pointing left & right, 
"++" for a symmetric tie, and ":" for sets of vertices

```{r}
plot(graph_from_literal(a---b, b---c)) # the number of dashes doesn't matter
plot(graph_from_literal(a--+b, b+--c))
plot(graph_from_literal(a+-+b, b+-+c)) 
plot(graph_from_literal(a:b:c---c:d:e))

gl <- graph_from_literal(a-b-c-d-e-f, a-g-h-b, h-e:f:i, j)
plot(gl)
```


## Edge, vertex, and network attributes

Access vertices and edges:

```{r}
E(g4) # The edges of the object
V(g4) # The vertices of the object
```

You can also manipulate the network matrix directly:

```{r}
g4[]
g4[1,]
g4[3,3] <- 10
g4[5,7] <- 10
```

Add attributes to the network, vertices, or edges:

```{r}
V(g4)$name # automatically generated when we created the network.
V(g4)$gender <- c("male", "male", "male", "male", "female", "female", "male")
E(g4)$type <- "email" # Edge attribute, assign "email" to all edges
E(g4)$weight <- 10    # Edge weight, setting all existing edges to 10
```

Examine attributes

```{r}
edge_attr(g4)
vertex_attr(g4)
graph_attr(g4)
```

Another way to set attributes
(you can similarly use set_edge_attr(), set_vertex_attr(), etc.)

```{r}
g4 <- set_graph_attr(g4, "name", "Email Network")
g4 <- set_graph_attr(g4, "something", "A thing")

graph_attr_names(g4)
graph_attr(g4, "name")
graph_attr(g4)

g4 <- delete_graph_attr(g4, "something")
graph_attr(g4)
```

```{r}
plot(g4, edge.arrow.size=.5, vertex.label.color="black", vertex.label.dist=1.5,
     vertex.color=c( "pink", "skyblue")[1+(V(g4)$gender=="male")] ) 
```

g4 has two edges going from Jim to Jack, and a loop from John to himself.
We can simplify our graph to remove loops & multiple edges between the same nodes.
Use 'edge.attr.comb' to indicate how edge attributes are to be combined - possible 
options include "sum", "mean", "prod" (product), min, max, first/last (selects 
the first/last edge's attribute). Option "ignore" says the attribute should be 
disregarded and dropped.

```{r}
g4s <- simplify( g4, remove.multiple = T, remove.loops = F, 
                 edge.attr.comb=list(weight="sum", type="ignore") )
plot(g4s, vertex.label.dist=1.5)
g4s
```

Let's take a look at the description of the igraph object.
Those will typically start with up to four letters:
1. D or U, for a directed or undirected graph
2. N for a named graph (where nodes have a name attribute)
3. W for a weighted graph (where edges have a weight attribute)
4. B for a bipartite (two-mode) graph (where nodes have a type attribute)

The two numbers that follow refer to the number of nodes and edges in the graph. 
The description also lists graph, node & edge attributes, for example:
(g/c) - graph-level character attribute
(v/c) - vertex-level character attribute
(e/n) - edge-level numeric attribute


## Specific graphs and graph models

### Empty graph

```{r}
eg <- make_empty_graph(40)
plot(eg, vertex.size=10, vertex.label=NA)
```

### Full graph

```{r}
fg <- make_full_graph(40)
plot(fg, vertex.size=10, vertex.label=NA)
```

### Star graph 

```{r}
st <- make_star(40)
plot(st, vertex.size=10, vertex.label=NA) 
```

### Tree graph


```{r}
tr <- make_tree(40, children = 3, mode = "undirected")
plot(tr, vertex.size=10, vertex.label=NA) 
```

### Ring graph


```{r}
rn <- make_ring(40)
plot(rn, vertex.size=10, vertex.label=NA)
```

### Erdos-Renyi random graph 

('n' is number of nodes, 'm' is the number of edges)

```{r}
er <- sample_gnm(n=100, m=40) 
plot(er, vertex.size=6, vertex.label=NA)  
```

### Watts-Strogatz small-world graph
Creates a lattice with 'dim' dimensions of 'size' nodes each, and rewires edges 
randomly with probability 'p'. You can allow 'loops' and 'multiple' edges.
The neighborhood in which edges are connected is 'nei'.

```{r}
sw <- sample_smallworld(dim=2, size=10, nei=1, p=0.1)
plot(sw, vertex.size=6, vertex.label=NA, layout=layout_in_circle)
```
 
### Barabasi-Albert preferential attachment model for scale-free graphs
'n' is number of nodes, 'power' is the power of attachment (1 is linear)
'm' is the number of edges added on each time step 

```{r}
 ba <-  sample_pa(n=100, power=1, m=1,  directed=F)
 plot(ba, vertex.size=6, vertex.label=NA)
```


# Case Study: The Karate Club
 
Let's study the Zachary Karate Club. This is a social network between members of a university karate club, led by president John A. and karate instructor Mr. Hi (pseudonyms).

The edge weights are the number of common activities the club members took part of. These
activities were:
1. Association in and between academic classes at the university.
2. Membership in Mr. Hi’s private karate studio on the east side of the city where Mr. Hi taught
nights as a part-time instructor.
3. Membership in Mr. Hi’s private karate studio on the east side of the city, where many of his
supporters worked out on weekends.
4. Student teaching at the east-side karate studio referred to in (2). This is different from (2)
in that student teachers interacted with each other, but were prohibited from interacting with
their students.
5. Interaction at the university rathskeller, located in the same basement as the karate club’s
workout area.
6. Interaction at a student-oriented bar located across the street from the university campus.
7. Attendance at open karate tournaments held through the area at private karate studios.
8. Attendance at intercollegiate karate tournaments held at local universities. Since both open
and intercollegiate tournaments were held on Saturdays, attendance at both was impossible.


Zachary studied conflict and fission in this network, as the karate club was split into two separate
clubs, after long disputes between two factions of the club, one led by John A., the other by Mr. Hi.
The ‘Faction’ vertex attribute gives the faction memberships of the actors. After the split of the
club, club members chose their new clubs based on their factions, except actor no. 9, who was in
John A.’s faction but chose Mr. Hi’s club.


```{r}
g <- graph.ring(10,directed=TRUE)               
plot(g)
ShortPth <- get.shortest.paths(g, 8, 2)    # List of path 8->2
ShortPth
E(g)$color <- "SkyBlue2"
E(g)$width <- 1
E(g, path=ShortPth$vpath[[1]])$color <- "red" 

### setting edges by path= is failing !!!!!!!!!!
plot(g)

```

```{r}
data("karate")
karate
summary(karate)
head(karate)
karate[1,]

l <- layout_with_fr(karate)
plot(karate, vertex.size=17, edge.width=2, layout=l)
```

```{r}
deg <- degree(karate, mode="all")
deg


centr_clo(karate, mode="all", normalized=T) 


centr_betw(karate, directed=T, normalized=T)

hist(deg, breaks=1:vcount(karate)-1, main="Histogram of node degree")

plot(karate, vertex.size=deg*2)
```
 
## Rewiring a graph

'each_edge()' is a rewiring method that changes the edge endpoints
uniformly randomly with a probability 'prob'.

```{r}
 rn.rewired <- rewire(rn, each_edge(prob=0.1))
 plot(rn.rewired, vertex.size=10, vertex.label=NA)
 
 # Rewire to connect vertices to other vertices at a certain distance. 
 rn.neigh = connect.neighborhood(rn, 5)
 plot(rn.neigh, vertex.size=8, vertex.label=NA) 
 
 
 # Combine graphs (disjoint union, assuming separate vertex sets): %du%
 plot(rn, vertex.size=10, vertex.label=NA) 
 plot(tr, vertex.size=10, vertex.label=NA) 
 plot(rn %du% tr, vertex.size=10, vertex.label=NA) 
```

  
 
# ================ 3. Reading network data from files ================

Set the working directory to the folder containing the workshop files:

```{r}
setwd("~/sandbox/NetSciX Workshop")  
```


## DATASET 1: edgelist 

```{r}
nodes <- read.csv("Dataset1-Media-Example-NODES.csv", header=T, as.is=T)
links <- read.csv("Dataset1-Media-Example-EDGES.csv", header=T, as.is=T)
```

Examine the data:

```{r}
head(nodes)
head(links)
nrow(nodes); length(unique(nodes$id))
nrow(links); nrow(unique(links[,c("from", "to")]))
```

Collapse multiple links of the same type between the same two nodes
by summing their weights, using aggregate() by "from", "to", & "type":
(we don't use "simplify()" here so as not to collapse different link types)

```{r}
links <- aggregate(links[,3], links[,-3], sum)
links <- links[order(links$from, links$to),]
colnames(links)[4] <- "weight"
rownames(links) <- NULL
```


## DATASET 2: matrix 


```{r}
nodes2 <- read.csv("Dataset2-Media-User-Example-NODES.csv", header=T, as.is=T)
links2 <- read.csv("Dataset2-Media-User-Example-EDGES.csv", header=T, row.names=1)
```

Examine the data:

```{r}
head(nodes2)
head(links2)

links2 is an adjacency matrix for a two-mode network:
links2 <- as.matrix(links2)
dim(links2)
dim(nodes2)
```


# Turning networks into igraph objects
 


## DATASET 1

Converting the data to an igraph object:
The graph.data.frame function, which takes two data frames: 'd' and 'vertices'.
'd' describes the edges of the network - it should start with two columns 
containing the source and target node IDs for each network tie.
'vertices' should start with a column of node IDs.
Any additional columns in either data frame are interpreted as attributes.

```{r}
net <- graph_from_data_frame(d=links, vertices=nodes, directed=T) 
```

Examine the resulting object:

```{r}
class(net)
net 
```

 We can look at the nodes, edges, and their attributes:
 
```{r}
E(net)
V(net)
E(net)$type
V(net)$media
```


```{r}
plot(net, edge.arrow.size=.4,vertex.label=NA)
```

Removing loops from the graph:

```{r}
net <- simplify(net, remove.multiple = F, remove.loops = T) 
```

If you need them, you can extract an edge list or a matrix from igraph networks.

```{r}
as_edgelist(net, names=T)
as_adjacency_matrix(net, attr="weight")

# Or data frames describing nodes and edges:
as_data_frame(net, what="edges")
as_data_frame(net, what="vertices")
```

## DATASET 2

```{r}
head(nodes2)
head(links2)

net2 <- graph_from_incidence_matrix(links2)

# A built-in vertex attribute 'type' shows which mode vertices belong to.
table(V(net2)$type)

plot(net2,vertex.label=NA)

# To transform a one-mode network matrix into an igraph object,
# use graph_from_adjacency_matrix()

# We can also easily generate bipartite projections for the two-mode network:
# (co-memberships are easy to calculate by multiplying the network matrix by
# its transposed matrix, or using igraph's bipartite.projection function)

net2.bp <- bipartite.projection(net2)

# We can calculate the projections manually as well:
#   as_incidence_matrix(net2)  %*% t(as_incidence_matrix(net2))
# t(as_incidence_matrix(net2)) %*%   as_incidence_matrix(net2)

plot(net2.bp$proj1, vertex.label.color="black", vertex.label.dist=1,
     vertex.label=nodes2$media[!is.na(nodes2$media.type)])

plot(net2.bp$proj2, vertex.label.color="black", vertex.label.dist=1,
     vertex.label=nodes2$media[ is.na(nodes2$media.type)])
```

