---
title: "SQL Challenge"
---


```{eval=FALSE, include=FALSE}

The schema for the tables:

create table master (
playerID text,
birthYear integer,
birthMonth integer,
birthDay integer,
birthCountry text,
birthState text,
birthCity text,
deathYear integer,
deathMonth int,
deathDay int,
deathCountry text,
deathState text,
deathCity text,
nameFirst text,
nameLast text,
nameGiven text,
weight int,
height int,
bats  text,
throws text,
debut date,
finalGame  text,
retroID text,
bbrefID text,
deathDate date,
birthDate date
);

create table salaries(
yearID int,
teamID text,
lgID text,
playerID text,
salary int
)
```

```{r, message=FALSE}
library(tidyverse) 
library(sqldf)
library(Lahman)

str(Master)
str(Salaries)

#write_csv(Salaries, path="Salaries.csv")
 
head(Master, n=10)
```


### Let's just look at the structure of the tables we will use


```{r}
sqldf("select * from Master limit 10")
```


# Easy Questions

## Question: When and where was Derek Jeter born?

```{r}
sqldf("SELECT birthDate, birthCity, birthState, birthCountry from Master where nameFirst=\"Derek\" and nameLast=\"Jeter\"")
```


## Question: How many unique players are there in the database?

```{r}
sqldf("SELECT count(distinct playerID) from Master")
```

## Question: What is the highest salary ever?

```{r}
sqldf("SELECT max(salary) from Salaries")
```

## Question: What is the average salary for each year?

```{r}
sqldf("SELECT yearID, AVG(salary) FROM Salaries GROUP BY yearID ORDER BY yearID ASC")
```



## Question: How much did Derek Jeter (playerID=jeterde01) make throughout his career?


```{r}
sqldf("SELECT yearID, salary FROM Salaries where playerID=\"jeterde01\"")
```

## Question: How many players were born in the 80s?

```{r}
sqldf("SELECT COUNT(*) from Master WHERE birthYear >= 1980 AND birthYear <= 1989")
```



# Medium Questions

## Questions: What are the top 5 the most popular birth states in the USA?

```{r}
sqldf("select birthState, count(*) as stateCount FROM Master WHERE birthCountry=\"USA\" GROUP BY birthState ORDER BY stateCount DESC limit 5")
```

## Question: What is the average salary for each birth year?

```{r}
sqldf("SELECT Master.birthYear, count(*) as NumPlayers, AVG(Salaries.salary) FROM Salaries JOIN Master ON Salaries.playerID = Master.playerID  GROUP BY Master.birthYear ORDER BY Master.birthYear ASC")
```




## Question: Who has the highest average salary throughout their career?

```{r}
sqldf("select playerID, count(*) as num, avg(salary) as avgSalary from Salaries group by playerID order by avgSalary desc")
```





# Hard Questions

## Question: Which player, and who did he play for, had the lowest salary in 1994?

```{r}
sqldf("select Master.nameGiven, Master.playerID, Salaries.salary, Salaries.yearID, Salaries.teamID FROM Salaries JOIN Master ON Salaries.playerID = Master.playerID WHERE Salaries.yearID = 1994 ORDER BY Salaries.salary limit 10")
```


## Question: 


